A survey of generic IND-CCA2 transformations

### 2024-07-07 faster-fo-transform/cca-transform-survey/main.tex
Checkpoint

\begin{definition}[$\delta$-correctness]\label{def:delta-correctness}
    A probabilistic public-key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$ is $\delta$-correct if the expected maximal probability of decryption failure taken across the distribution of keypairs is at most $\delta$:

@@

\begin{table}[H]
    \center
    \begin{tabular}{c|c|c}
        \monospace{ATK} & $\mathcal{O}_1$ & $\mathcal{O}_2$ \\
        \hline
        \monospace{CPA} & \multicolumn{2}{|c|}{$-$} \\
        \hline
        \monospace{CCA1} & $\mathcal{O}^\decrypt$ & - \\
        \hline
        \monospace{CCA2} & \multicolumn{2}{|c|}{$\mathcal{O}^\decrypt$} \\
        \hline
        \monospace{PCVA} & \multicolumn{2}{|c|}{$\pco, \cvo$} \\
        \hline
        \monospace{PCA} & \multicolumn{2}{|c|}{$\pco$} \\
        \hline
        \monospace{VA} & \multicolumn{2}{|c|}{$\cvo$} \\
    \end{tabular}
    \caption{Attacks and associated oracle access}
\end{table}

@@

\section{Modular Fujisaki-Okamoto transformation}
The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} and its KEM variations \cite{hofheinz2017modular} achieve security against adaptive chosen-ciphertext attacks through \emph{de-randomization} and \emph{re-encryption}. In particular, the modular FO transformation contains two steps. First, the

>>>>

\begin{definition}[$\delta$-correctness]\label{def:delta-correctness-pke}
    A probabilistic public-key encryption scheme $\pke = (\keygen, \encrypt, \decrypt)$ is $\delta$-correct if the expected maximal probability of decryption failure taken across the distribution of keypairs is at most $\delta$:

@@

\begin{table}[H]
    \center
    \begin{tabular}{|c|c|c|}
        \hline
        \monospace{ATK} & $\mathcal{O}_1$ & $\mathcal{O}_2$ \\
        \hline
        \monospace{CPA} & \multicolumn{2}{|c|}{$-$} \\
        \hline
        \monospace{CCA1} & $\mathcal{O}^\decrypt$ & - \\
        \hline
        \monospace{CCA2} & \multicolumn{2}{|c|}{$\mathcal{O}^\decrypt$} \\
        \hline
        \monospace{PCVA} & \multicolumn{2}{|c|}{$\pco, \cvo$} \\
        \hline
        \monospace{PCA} & \multicolumn{2}{|c|}{$\pco$} \\
        \hline
        \monospace{VA} & \multicolumn{2}{|c|}{$\cvo$} \\
        \hline
    \end{tabular}
    \caption{Attacks and associated oracle access}
\end{table}

@@

\subsection{Key encapsulation mechanism}
A key encapsulation mechanism $\kem = (\keygen, \encap, \decap)$ is a collection of three routines. The key generation routine $(\pk, \sk) \leftsample \keygen(1^\lambda)$ takes the security parameter $1^\lambda$ and returns a keypair. The encapsulation routine $(c, K) \leftsample \encap(\pk)$ takes the public key and outputs some ciphertext $c \in \mathcal{C}_\kem$ and some shared secret $K \in \mathcal{K}_\kem$. Finally, the decapsulation routine $K \leftarrow \decap(\sk, c)$ takes the secret key and a ciphertext and outputs the correponding shared secret.

\emph{Correctness}: similar to a $\pke$, key encapsulation mechanisms are usually required to be perfectly correct, meaning that for all keypairs $(\pk, \sk)$, decapsulation always outputs the same shared secret as the encapsulation

\begin{equation*}
    \Prob{(c, K_1) \leftsample \encap(\pk); K_2 \leftarrow \decap(\sk, c); K_1 = K_2} = 1
\end{equation*}

However, where decapsulation failure has a non-zero probability, we need to upperbound this quantity.

\begin{definition}[$\delta$-correctness]\label{def:delta-correctness-kem}
    A $\kem = (\keygen, \encap, \decap)$ is $\delta$-correct if the probability of decapsulation failure taken across the keypair distribution is at most $\delta$:

    \begin{equation*}
        P\lrbrack{
            \decap(\sk, c) \neq K 
            \mid (\pk, \sk) \leftsample \keygen()
                ; (c, K) \leftsample \encap(\pk)
        } \leq \delta
    \end{equation*}
\end{definition}

\emph{Security}: the security of a $\kem$ is defined in an adversarial game in which the adversary's goal is to distinguish between shared secret derived from encapsulation and uniformly random samples. An adversary $A = (A_1, A_2)$ contains two sub-routines with access to some oracle $\mathcal{O}$ depending on game.

\begin{figure}[H]
    \center
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{\monospace{IND-CCA2} game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $s \leftsample A_1^{\mathcal{O}^\decap}(1^\lambda, \pk)$
                \State $(c^\ast, K_0) \leftsample \encap(\pk)$
                \State $K_1 \leftsample \mathcal{K}_\kem$
                \State $b \leftsample \{0, 1\}$
                \State $\hat{b} \leftsample A_2^{\mathcal{O}^\decap}(1^\lambda, \pk, s, c^\ast, K_b)$
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\decap$ oracle $\mathcal{O}^\decap(c \neq c^\ast)$}
            \begin{algorithmic}[1]
                \State \Return $\decap(\sk, c)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The \monospace{IND-CCA2} game for $\kem$}\label{fig:kem-ind-cca2-game}
\end{figure}

\begin{definition}[\monospace{IND-CCA2 security}]\label{def:kem-ind-cca2-security}
    A $\kem = (\keygen, \encap, \decap)$ is \monospace{IND-CCA2} secure if no efficient adversary has non-negligible advantage in the \monospace{IND-CCA2} game.
\end{definition}

\section{Modular Fujisaki-Okamoto transformation}
The Fujisaki-Okamoto transformation \cite{fujisaki1999secure} and its KEM variations \cite{hofheinz2017modular} achieve security against adaptive chosen-ciphertext attacks through \emph{de-randomization} and \emph{re-encryption}. In particular, the modular FO transformation contains two steps. The first step (denoted the $T$ transformation) takes a $\pke = (\keygen, \encrypt, \decrypt)$ and outputs a $\pke_1 = (\keygen, \encrypt_1, \decrypt_1)$. The key generation remains unchanged, but the encryption routine is \emph{de-randomized} by deriving the coin from the plaintext using a hash function $G: \mathcal{M}_\pke \rightarrow \mathcal{R}_\pke$, and the decryption routine uses \emph{re-encryption} to reject invalid ciphertexts.

\begin{figure}[H]
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\encrypt_1(\pk, m)$}\label{alg:encrypt-1}
            \begin{algorithmic}[1]
                \State $r \leftarrow G(m)$
                \State $c \leftarrow \encrypt(\pk, m; r)$
                \State \Return $c$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \hfill
    \begin{minipage}{0.49\textwidth}
        \begin{algorithm}[H]
            \caption{$\decrypt_1(\sk, c)$}\label{alg:decrypt-1}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow D(\sk, c)$
                \If{
                    $\hat{m} \in \mathcal{M}_\pke 
                    \;\land\; \encrypt(\pk, \hat{m}; G(\hat{m})) = c$
                }
                    \State \Return $\hat{m}$
                \EndIf
                \State \Return $\bot$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{T transformation}\label{fig:t-transformation}
\end{figure}

\cite{hofheinz2017modular} states the security property of $\pke_1$ as follows:

\begin{theorem}\label{thm:owcpa-implies-owpcva}
    If $\pke$ is $\delta$-correct and has $\gamma$ spread, then for every \monospace{OW-PCVA} adversary $B$ against $\pke_1$ who makes $q_G$ hash queries, $q_V$ ciphertext validation queries, and $q_P$ plaintext checking queries, there exists an \monospace{OW-CPA} adversary $A$ such that:

    \begin{equation*}
        \monospace{Adv}(B) \leq q_V \cdot 2^{-\gamma}
            + (q_G + q_P) \cdot \delta
            + (1 + q_G + q_P) \cdot \monospace{Adv}(A)
    \end{equation*}
\end{theorem}

In other words, if $\pke$ is \monospace{OW-CPA} secure, then $\pke_1$ is \monospace{OW-PCVA} secure, though the security is non-tight. On the other hand, if $\pke$ is \monospace{IND-CPA} secure, the \monospace{OW-PCVA} security is tight:

\begin{theorem}\label{thm:indcpa-implies-owpcva}
    for every \monospace{OW-PCVA} adversary $B$ against $\pke_1$ who makes $q_G$ hash queries, $q_V$ ciphertext validation queries, and $q_P$ plaintext checking queries, there exists an \monospace{IND-CPA} adversary $A$ such that:

    \begin{equation*}
        \monospace{Adv}(B) \leq q_V \cdot 2^{-\gamma}
            + (q_G + q_P) \cdot \delta
            + \frac{2 \cdot q_G + 1}{\norm{\mathcal{M}_\pke}}
            + 3 \cdot \monospace{Adv}(A)
    \end{equation*}
\end{theorem}

\begin{proof}
    We will provide a sketch of proof for theorem \ref{thm:owcpa-implies-owpcva} and \ref{thm:indcpa-implies-owpcva}. The main idea is to construct an \monospace{OW-CPA}/\monospace{IND-CPA} adversary $A$ who can simulate the \monospace{OW-PCVA} game and use the \monospace{OW-PCVA} adversary $B$ as a sub-routine. However, there are three main difficulties with constructing a ``convincing'' simulation:

    \begin{enumerate}
        \item $A$ has no access to $\pco$
        \item $A$ has no access to $\cvo$
        \item The challenge ciphertext $A$ receives is obtained using a truly random coin, but the challenge ciphertext $B$ expects is obtained using a pseudorandom coin $r \leftarrow G(m)$
    \end{enumerate}
\end{proof}
